Bayesian multivariate skew-normal finite mixture model for analysis of infant development trajectories

### 2019-06-04 manuscript/manuscript.tex
refs
mods

@@

The general formulation of the multivariate skew normal (MSN) is as follows.

@@

@@

We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model. We adopt a standard stochastic representation of a MSN random variable as the sum of a latent truncated normal random effect and a correlated multivariate normal random error term. More formally, Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$. The MSN stochastic representation is as follows.
$$\mathbf{y}_{i} = \boldsymbol\beta^T \mathbf{x}_{i} + t_{i}\boldsymbol\psi + \boldsymbol\epsilon_{i}$$

@@

>>>>

\subsection{Infant Development Clustering}

Heterogeneity of treatment effects (HTE) (Lanza and Rhoades, 2013).

@@

Let $\mathbf{y}_{i}$ be the $J \times 1$ observation vector for subject $i$ such that $y_{ij}$ is the observation for subject $i$ at timepoint $j$. The general formulation of the multivariate skew normal mixure model is as follows.

@@

Here $\boldsymbol\theta_k$ is the set of parameters shared by subjects belonging to class $k$ ($k = 1,...,K$). The probability of subject $i$ ($i = 1,...,n$) belonging to class $k$ ($k = 1,...,K$) is denoted $\pi_{ik}$. Note that $\pi_{ik}$ is not indexed by time and we do not allow for subjects switching classes over the study period. This feature of the model is a result of our interest in clustering based on development patterns rather than time-specific development scores.

In each development class, we assume that a subject's $J\times 1$ outcome vector follows a multivariate skew normal (MSN) distribution. We define a latent variable $Z_i \in \{1,...,K\}$ to denote the cluster membership of subject $i$ for $i = 1,...,n$. We model $\mathbf{y}_{i}|Z_i$ as

$$f(\mathbf{y}_{i_k}|Z_i )$$

@@

We model the effect of covariates on longitudinal development outcomes through the use of a MSN regression model. We adopt a standard stochastic representation of a MSN random variable as the sum of a latent truncated normal random effect and a correlated multivariate normal random error term. The MSN stochastic representation is as follows.
$$\mathbf{y}_{i} = \boldsymbol\beta^T \mathbf{x}_{i} + t_{i}\boldsymbol\psi + \boldsymbol\epsilon_{i}$$

@@

\bibitem{ } Lanza ST, Rhoades BL. Latent class analysis: an alternative perspective on subgroup analysis in prevention and treatment. \textit{Prevention Science}. 2013 Apr 1;14(2):157-68.
